---
title: "CensusAggregator Report"
output:
  html_document:
    css: "style.css"
    theme:
      version: 4
params:
  geo_uid: !r c("5750001.00")
  geography: "ct"
  bookmark: "https://censusaggregator.ca/"
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE, error = FALSE
)
```

```{r html-utils}
section_heading <- function(title) {
  htmltools::div(
    htmltools::hr(),
    htmltools::div(
      class = "section-heading",
      htmltools::h1(title),
      htmltools::hr()
    )
  )
}

subsection_heading <- function(title, class = NULL) {
  htmltools::div(
    class = "subsection-heading",
    htmltools::h2(title)
  )
}

snapshot_card_value <- function(value) {
  htmltools::div(
    class = "card-value",
    value
  )
}

snapshot_card <- function(title, value, ..., class = NULL) {
  htmltools::div(
    class = paste("box", class),
    htmltools::div(htmltools::HTML(title)),
    snapshot_card_value(value),
    htmltools::div(...)
  )
}

header_and_barchart <- function(...) {
  shiny::column(
    width = 6,
    ...
  )
}

pull_and_format_value <- function(data, format = "percent", column = "value") {
  value <- data[[column]]

  if (is.na(value)) {
    return(htmltools::HTML("&#8212;"))
  }

  switch(format,
    percent = scales::percent(value, 0.1),
    comma = scales::comma(value, 1),
    comma_decimal = scales::comma(value, 0.1),
    dollar = scales::dollar(value)
  )
}

logo <- function(pagebreak = TRUE) {
  shiny::div(
    class = "print-header",
    shiny::div(
      class = "left-logo",
      shiny::img(
        src = system.file("app/www/logo.png", package = "censusaggregatorapp"),
        alt = "CensusAggregator logo"
      )
    ),
    shiny::div(
      class = "right-logo",
      shiny::img(
        src = system.file("app/www/pa-logo.png", package = "censusaggregatorapp"),
        alt = "Purpose Analytics logo"
      )
    )
  )
}

suppressed_note <- function(...) {
  data <- list(...)

  if (any(data == "&#8212;")) {
    shiny::div(
      shiny::HTML("<i>Note: &#8212; indicates data for one or more of the selected areas is not available or is suppressed due to confidentiality.</i>")
    )
  }
}
```

```{r packages}
library(censusaggregate)
library(dplyr)
library(purrr)
library(scales)
library(tidyr)
library(stringr)
library(bslib)
library(arrow)
library(censusaggregatorapp)
library(sfarrow)
library(ggplot2)
library(sf)
library(forcats)
```

```{r data}
geography <- params$geography
regions <- params$geo_uid

data <- prepare_data(geography, regions) %>%
  split(.$label) %>%
  purrr::map(~ .x %>%
    select(-label) %>%
    rename(label = breakdown))
```

```{r}
logo(FALSE)
```

```{r map, fig.height = 3.25, dpi = 150, fig.alt = "A map showing the outline of the selected geographies."}
boundaries_dataset <- glue::glue("extdata/{geography}")

boundaries_data <- open_dataset(system.file(boundaries_dataset, package = "censusaggregatorapp"))
boundaries_data <- boundaries_data %>%
  read_sf_dataset()

# Limit to regions, combine boundaries
region_boundaries <- boundaries_data %>%
  filter(geo_uid %in% regions)
region_boundaries <- region_boundaries %>%
  st_union()

# Get bounding box
regions_bbox <- region_boundaries %>%
  st_bbox()

# Add 10% buffer to bbox, based on which of height/width is smaller
smaller_size <- min(regions_bbox$xmax - regions_bbox$xmin, regions_bbox$ymax - regions_bbox$ymin)
bbox_buffer <- smaller_size * 0.05

regions_bbox[["xmin"]] <- regions_bbox$xmin - bbox_buffer
regions_bbox[["ymin"]] <- regions_bbox$ymin - bbox_buffer
regions_bbox[["xmax"]] <- regions_bbox$xmax + bbox_buffer
regions_bbox[["ymax"]] <- regions_bbox$ymax + bbox_buffer

# Get all boundaries within bbox
all_boundaries_in_bbox <- boundaries_data %>%
  st_crop(regions_bbox)

ggplot() +
  geom_sf(data = all_boundaries_in_bbox, colour = "grey", fill = NA, size = 0.3) +
  geom_sf(data = region_boundaries, colour = "black", fill = "white", alpha = 0.75, size = 0.5) +
  geom_rect(aes(xmin = regions_bbox$xmin, xmax = regions_bbox$xmax, ymin = regions_bbox$ymin, ymax = regions_bbox$ymax), fill = NA, colour = "grey", size = 0.3) +
  theme_void()

# tmap::tm_shape(all_boundaries_in_bbox) + tmap::tm_polygons(alpha = 0, border.col = "grey") + tmap::tm_shape(region_boundaries) + tmap::tm_polygons(alpha = 0.75, border.col = "black", col = "white") + tmap::tm_layout(frame = FALSE)

# plot(all_boundaries_in_bbox[["geometry"]], border = "grey", xlim = c(regions_bbox$xmin, regions_bbox$xmax), ylim = c(regions_bbox$ymin, regions_bbox$ymax))
# par(new = TRUE, bg = NA)
# plot(region_boundaries, col = scales::alpha("white", 0.4), xlim = c(regions_bbox$xmin, regions_bbox$xmax), ylim = c(regions_bbox$ymin, regions_bbox$ymax))
```

```{r}
shiny::div(
  class = "censusagg-link",
  shiny::tags$a(href = params$bookmark, "Open in CensusAggregator", target = "_blank")
)
```

::: {.row .mb-2 .boxes}

```{r snapshot}
population <- data[["Population, 2021"]] %>%
  pull_and_format_value("comma")

households <- data[["Households"]] %>%
  pull_and_format_value("comma")

population_change <- data[["Population change, 2016 to 2021"]] %>%
  pull_and_format_value()

if (!str_starts(population_change, "-")) {
  population_change <- paste0("+", population_change)
}

population_density <- data[["Population density"]] %>%
  pull_and_format_value("comma_decimal")
```

```{r population}
shiny::column(
  width = 3,
  snapshot_card("Population", population)
)
```

```{r households}
shiny::column(
  width = 3,
  snapshot_card("Households", households)
)
```

```{r population_change}
shiny::column(
  width = 3,
  snapshot_card("Population change<br>(2016 to 2021)", population_change)
)
```

```{r population_density}
shiny::column(
  width = 3,
  snapshot_card("Population density<br>(people/km<sup>2</sup>)", population_density)
)
```
:::

```{r}
suppressed_note(population, households, population_change, population_density)
```

```{r section-age-households-families}
section_heading("Age, households, and families")
```

::: {.row .mb-2}

```{r age-cohorts-household-size-family-type}
header_and_barchart(
  subsection_heading("Population by age cohorts"),
  data[["Age (cohorts)"]] %>%
    inline_barchart(),
  subsection_heading("Private households by size"),
  data[["Household size"]] %>%
    inline_barchart()
)
```

```{r household_type}
header_and_barchart(
  subsection_heading("Family type"),
  data[["Family type"]] %>%
    inline_barchart(),
  subsection_heading("Household type"),
  data[["Household type"]] %>%
    inline_barchart()
)
```

:::

<div class = "pagebreak"></div>

```{r section-language}
section_heading("Language")
```

::: {.row .mb-2 .boxes}

```{r knowledge_of_english}
knowledge_of_english <- data[["Knowledge of official languages"]] %>%
  filter(label == "Knowledge of English") %>%
  pull_and_format_value(column = "value_proportion")

shiny::column(
  width = 3,
  snapshot_card(
    "Knowledge of English", knowledge_of_english
  )
)
```

```{r knowledge_of_french}
knowledge_of_french <- data[["Knowledge of official languages"]] %>%
  filter(label == "Knowledge of French") %>%
  pull_and_format_value(column = "value_proportion")
shiny::column(
  width = 3,
  snapshot_card("Knowledge of French", knowledge_of_french)
)
```

```{r english-at-home}
english_at_home <- data[["Official languages most often spoken at home"]] %>%
  filter(label == "English") %>%
  pull_and_format_value(column = "value_proportion")
shiny::column(
  width = 3,
  snapshot_card("English spoken most often at home", english_at_home)
)
```

```{r french-at-home}
french_at_home <- data[["Official languages most often spoken at home"]] %>%
  filter(label == "French") %>%
  pull_and_format_value(column = "value_proportion")
shiny::column(
  width = 3,
  snapshot_card("French spoken most often at home", french_at_home)
)
```

:::

```{r}
suppressed_note(knowledge_of_english, knowledge_of_french, english_at_home, french_at_home)
```

::: {.row}

```{r language_at_home}
top_languages <- data[["Top non-official languages spoken most often at home"]]
if (is.null(top_languages)) {
  top_languages <- "No data on non-official languages spoken most often at home."
} else {
  top_languages <- data[["Top non-official languages spoken most often at home"]] %>%
    mutate(label = forcats::fct_inorder(label)) %>%
    inline_barchart()
}

header_and_barchart(
  subsection_heading("Top non-official languages spoken most often at home"),
  top_languages
)
```

:::

```{r section-income-housing}
section_heading("Income and housing")
```

::: {.row .mb-2 .boxes}

```{r median_income}
median_income_label <- ifelse(length(regions) == 1, "Median household income", "Estimated median household income")

median_income <- data[[median_income_label]] %>%
  pull_and_format_value("dollar")

shiny::column(
  width = 4,
  snapshot_card(median_income_label, median_income)
)
```

```{r lim_at}
lim_at <- data[["Low-income measure after tax (LIM-AT)"]] %>% pull_and_format_value(column = "value_proportion")
shiny::column(
  width = 4,
  snapshot_card("Low income (LIM-AT)", lim_at)
)
```

```{r unaffordable_housing}
unaffordable_housing <- data[["Unaffordable housing"]] %>% pull_and_format_value(column = "value_proportion")
shiny::column(
  width = 4,
  snapshot_card("Unaffordable housing", unaffordable_housing)
)
```

:::

```{r}
suppressed_note(median_income, lim_at, unaffordable_housing)
```

::: {.row}

```{r total-income}
header_and_barchart(
  subsection_heading("Total household income"),
  data[["Total household income ($20,000 buckets)"]] %>%
    mutate(label = forcats::fct_inorder(label)) %>%
    inline_barchart()
)
```

```{r shelter-cost}
shelter_cost <- data[["Average shelter cost"]]

# If they are all NA, then fake as proportion so that no bars are shown
shelter_cost_format <- ifelse(all(is.na(shelter_cost[["value"]])), "proportion", "dollar")

# Remove 'count' field if still there, fix widths
shelter_cost <- shelter_cost %>%
  inline_barchart(shelter_cost_format)

if ("count" %in% names(shelter_cost[["_data"]])) {
  shelter_cost <- shelter_cost %>%
    gt::cols_hide(count) %>%
    gt::cols_width(label ~ 250, value_fmt ~
    75, hist ~ 100)
}

header_and_barchart(
  subsection_heading("Average shelter cost"),
  shelter_cost,
  subsection_heading("Household tenure"),
  data[["Tenure"]] %>%
    mutate(label = fct_inorder(label)) %>%
    inline_barchart()
)
```

:::

<div class = "pagebreak"></div>

```{r section-diversity-immigration}
section_heading("Diversity and immigration")
```

::: {.row .mb-2 .boxes}

```{r visible-minority-summary, eval = TRUE}
visible_minority_summary <- data[["Visible minority population"]] %>%
  filter(label == "Total visible minority population") %>%
  pull_and_format_value(column = "value_proportion")

shiny::column(
  width = 3,
  snapshot_card("Visible minority", visible_minority_summary)
)
```

```{r immigrant-status, eval = TRUE}
immigrant_status_total <- data[["Immigrant status"]] %>%
  filter(label == "Immigrants") %>%
  pull_and_format_value(column = "value_proportion")

immigrant_status_recent <- data[["Immigrant status"]] %>%
  filter(label == "Recent immigrants (2016 to 2021)") %>%
  pull_and_format_value(column = "value_proportion")

shiny::column(
  width = 3,
  snapshot_card("Total immigrants", immigrant_status_total)
)

shiny::column(
  width = 3,
  snapshot_card("Recent immigrants<br>(2016 to 2021)", immigrant_status_recent)
)
```

```{r indigenous-identity}
indigenous_identity <- data[["Indigenous identity"]] %>%
  pull_and_format_value(column = "value_proportion")
shiny::column(
  width = 3,
  snapshot_card(
    "Indigenous identity", indigenous_identity
  )
)
```

:::

```{r}
suppressed_note(visible_minority_summary, immigrant_status_total, immigrant_status_recent, indigenous_identity)
```

::: {.row}

```{r visible-minority-breakdown}
header_and_barchart(
  subsection_heading("Visible minority population"),
  data[["Visible minority population"]] %>%
    filter(label != "Total visible minority population") %>%
    derive_census_vector_order(by_value = TRUE) %>%
    inline_barchart()
)
```

```{r ethnic_or_cultural_origin, eval}
ethnic_or_cultural_origin <- data[["Top 10 ethnic or cultural origins"]]

if (is.null(ethnic_or_cultural_origin)) {
  ethnic_or_cultural_origin <- "No data available on ethnic or cultural origin."
} else {
  ethnic_or_cultural_origin <- data[["Top 10 ethnic or cultural origins"]] %>%
    mutate(label = forcats::fct_inorder(label)) %>%
    inline_barchart()
}

header_and_barchart(
  subsection_heading("Top ethnic or cultural origins"),
  ethnic_or_cultural_origin
)
```

:::

<hr>

<p>Data from CensusAggregator is adapted from Statistics Canada's 2021 Census of Population and may be freely used, reproduced, published, and distributed under the <a href = "https://www.statcan.gc.ca/eng/reference/licence" target = "_blank">Statistics Canada Open Data Licence</a>. Statistics Canada must be acknowledged as the source when using the data; acknowledgment of CensusAggregator is optional. The geographic boundaries have undergone some processing and may differ from the boundary files originally released by Statistics Canada. Statistics Canada randomly rounds to the nearest 0 or 5 - this can impact the precision of percentages shown in CensusAggregator.</p>

<div class = "pagebreak"></div>

```{r}
section_heading("Definitions")
```

::: {.row .definitions-table}

```{r variable-definitions}
# Split into two columns manually, after "Language spoken most often at home" will make it fit on 1 page

cutoff <- censusaggregatorapp::variable_definitions %>%
  mutate(id = row_number()) %>%
  filter(str_detect(variable_definition, "Language spoken most often at home")) %>%
  pull(id)

definitions_table <- function(data) {
  data %>%
    gt::gt() %>%
    gt::cols_align("left", variable_definition) %>%
    gt::fmt_markdown(columns = variable_definition) %>%
    gt::tab_options(
      column_labels.hidden = TRUE,
      table_body.border.bottom.color = "transparent",
      table_body.border.top.color = "transparent",
      table.font.names = "Lato"
    )
}

shiny::column(
  width = 6,
  censusaggregatorapp::variable_definitions %>%
    filter(row_number() <= cutoff) %>%
    definitions_table()
)
shiny::column(
  width = 6,
  censusaggregatorapp::variable_definitions %>%
    filter(row_number() > cutoff) %>%
    definitions_table()
)
```

:::
